---
title: 'Lab 01 - Functions and Vectors'
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    theme: cerulean
    highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F)
```



# Data

Data dictionary: [Syracuse city tax parcel data](https://ds4ps.github.io/Data-Science-Class/DATA/syr_parcels.html) 


```{r}
URL <- "https://raw.githubusercontent.com/DS4PS/Data-Science-Class/master/DATA/syr_parcels.csv"
dat <- read.csv( URL, stringsAsFactors=FALSE )
```

```{r, echo=F}
head( dat, 10 ) 
```

```{r, echo=F}
head( dat$owner, 10 )  # preview a vector
```



# Solutions


## Question 1

*How many tax parcels are in Syracuse (dataset dimensions)?* 

**ANSWER:** There are X number of tax parcels in Syracuse. 


```{r}
# your code here
```





## Question 2

*How many acres of land are in syracuse (sum over a numeric vector)?*

**ANSWER:** 





## Question 3

*How many vacant BUILDINGS are there in Syracuse (sum over a logical vector)?*

**ANSWER:** 



## Question 4

*What proportion of parcels are tax exempt?*

**ANSWER:** 




## Question 5

*Which neighborhood contains the most tax parcels (table with one variable)?*

**ANSWER:**



## Question 6

*Which neighborhood contains the most vacant LOTS (table with two variables)?*

**ANSWER:**



-----




----------- Delete text below this line before submission    -------------------


# Functions

You will use the following functions for this lab: 

```{r,eval=F}
names()                 # variable names
head()                  # preview dataset
length()                # vector dimensions
dim(), nrow(), ncol()   # dataset dimensions
sum(), summary()        # summarize numeric vectors
table()                 # summarize factors / character vectors
```

**Note:**

(1) Referencing variables in R requires both the **dataset name** and **variable name**, separated by the `$` operator:

```{r, eval=F}
summary( dat$acres )
```

(2) When you apply a **sum()** function to a numeric vector it returns the sum of all elements in the vector. 

```{r}
sum( c(10,20,5) )
```

When you apply a **sum()** function to a logical vector, it will count all of the TRUEs:

```{r}
x <- c(T,T,F,F,F)
sum( x )
sum( x ) / length( x ) # proportion of trues
```


(3) R wants to make sure you are aware of missing values, so it will return NA (not available) for functions performed on vectors with missing values. 

Add the 'NA remove' argument (`na.rm=TRUE`) to functions to ignore missing values: 

```{r, eval=F}
sum( dat$star, na.rm=T )
```






# Submission Instructions

When you have completed your assignment, knit your RMD file to generate your rendered HTML file. **Submit both your RMD and HTML files with solutions.**

Login to Canvas at <http://canvas.asu.edu> and navigate to the assignments tab in the course repository. Upload your solution files to the appropriate lab submission link.

Remember to:

* name your files according to the convention: **Lab-##-LastName.Rmd**
* show your solution - include your code.
* do not print excessive output (like a full data set or warning messages).
* follow appropriate style guidelines (spaces between arguments, etc.).

See [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml) for examples.

## Problems Uploading Files

Platforms like BlackBoard and Canvas may disallow you from submitting HTML files when there is embedded computer code (they sometimes think you are trying to upload a virus!). If this happens, create a zipped folder with both the RMD and HTML files and submit that instead.

## Notes on Knitting

Note that when you knit a file, it starts from a blank slate. You might have packages loaded or datasets active on your local machine, so you can run code chunks fine. But when you knit you might get errors that functions cannot be located or datasets don't exist. Be sure that you have included chunks to load these in your RMD file.

Your RMD file will not knit if you have errors in your code. If you get stuck on a question, just add `eval=F` to the code chunk and it will be ignored when you knit your file. That way I can give you credit for attempting the question and provide guidance on fixing the problem.





